Robinson et al. 2020 Supplemental Information: Comparing the Radiocarbon and Tree-ring Records for the Upland US Southwest

### 2020-03-11 analysis/paper/paper.Rmd
refining figures

# A smoothing distribution to use throughout the analyses
# This is a 21-year wide Gaussian distribution with a
# mean of 0 and standard deviation of 5 years.
# This is the smoother used in Bocinsky et al. 2016.
dist <- dnorm(seq(-10, 10, 1), sd = 5)

@@

dplyr::left_join(all_density,
  vep_demography,
  by = c(
    "Study Area",
    "Year AD"
  )
) %>%
  dplyr::group_by(
    `Study Area`,
    Type
  ) %>%
  dplyr::mutate(`Scaled Density` = Density * (sum(Population, na.rm = T) / sum(Density * !is.na(Population), na.rm = TRUE))) %>%
  dplyr::group_by(
    `Study Area`,
    Type,
    Period
  ) %>%
  dplyr::group_by(
    `Study Area`,
    Type
  ) %>%
  dplyr::mutate(
    `Raw Density` = ifelse(`Study Area` == "UUSW", Density, `Scaled Density`),
    `Smoothed Density` = ifelse(Type == "Tree-ring Density", stats::filter(`Raw Density`, filter = dist), `Raw Density`),
    `Raw Density` = ifelse(`Study Area` == "UUSW", `Raw Density`, NA)
  ) %>%
  dplyr::select(
    `Study Area`,
    Type,
    `Year AD`,
    Population,
    `Raw Density`,
    `Smoothed Density`,
    # `Binned Density`
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    Type = factor(Type,
      levels = c(
        "Tree-ring Density",
        "Radiocarbon Density",
        "VEP Population"
      ),
      ordered = TRUE
    ),
    `Study Area` = factor(`Study Area`,
      levels = c(
        `Upland US Southwest` = "UUSW",
        `Central Mesa Verde` = "CMV",
        `Northern Rio Grande` = "NRG"
      )
    )
  ) %>%
  ggplot(aes(x = `Year AD`)) +
  geom_line(aes(
    y = `Raw Density`,
    color = Type
  ),
  size = 0.5,
  alpha = 0.5
  ) +
  geom_line(aes(
    y = `Population`
  ),
  size = 1.25
  ) +
  geom_line(aes(
    y = `Smoothed Density`,
    color = Type
  ),
  size = 0.75
  ) +
  scale_color_manual(values = c(
    "Tree-ring Density" = "#d95f02",
    "Radiocarbon Density" = "#1b9e77",
    "VEP Population" = "black"
  )) +
  xlab("Year AD") +
  scale_x_continuous(
    limits = c(200, 1800),
    breaks = seq(200, 1800, 200),
    expand = expansion(0, 0)
  ) +
  theme_minimal(18) +
  theme(
    legend.title = element_blank(),
    legend.justification = c(0, 1),
    legend.position = "bottom",
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(size = 10),
    axis.title.y = element_blank(),
    # axis.text.y = element_blank(),
    # panel.grid.major.y = element_blank(),
    # panel.grid.minor.y = element_blank(),
    plot.margin = unit(c(0.25, 0.5, 0.25, 0.25), "in")
  ) +
  facet_grid(
    rows = vars(`Study Area`),
    scales = "free_y"
  )
```

>>>>

# This is a 21-year wide Gaussian distribution with a
# mean of 0 and standard deviation of 5 years.
# This is the smoother used in Bocinsky et al. 2016.
dist <- dnorm(seq(-10, 10, 1), sd = 5)

@@

dplyr::left_join(all_density,
  vep_demography,
  by = c(
    "Study Area",
    "Year AD"
  )
) %>%
  dplyr::group_by(
    `Study Area`,
    Type
  ) %>%
  dplyr::mutate(`Scaled Density` = Density * (sum(Population, na.rm = T) / sum(Density * !is.na(Population), na.rm = TRUE))) %>%
  dplyr::group_by(
    `Study Area`,
    Type,
    Period
  ) %>%
  dplyr::group_by(
    `Study Area`,
    Type
  ) %>%
  dplyr::mutate(
    `Raw Density` = ifelse(`Study Area` == "UUSW", Density, `Scaled Density`),
    `Smoothed Density` = ifelse(Type == "Tree-ring Density", stats::filter(`Raw Density`, filter = dist), `Raw Density`),
    `Raw Density` = ifelse(`Study Area` == "UUSW", `Raw Density`, NA)
  ) %>%
  dplyr::select(
    `Study Area`,
    Type,
    `Year AD`,
    Population,
    `Raw Density`,
    `Smoothed Density`,
    # `Binned Density`
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    Type = factor(Type,
      levels = c(
        "Tree-ring Density",
        "Radiocarbon Density",
        "VEP Population"
      ),
      ordered = TRUE
    ),
    `Study Area` = factor(`Study Area`,
      levels = c(
        "UUSW",
        "CMV",
        "NRG"
      ),
      labels = c(
        "Upland US Southwest Density",
        "Central Mesa Verde Population",
        "Northern Rio Grande Population"
      ),
      ordered = TRUE
    )
  ) %>%
  ggplot(aes(x = `Year AD`)) +
  geom_line(aes(
    y = `Raw Density`,
    color = Type
  ),
  size = 0.5,
  alpha = 0.5
  ) +
  geom_line(aes(
    y = `Population`
  ),
  size = 1.25
  ) +
  geom_line(aes(
    y = `Smoothed Density`,
    color = Type
  ),
  size = 0.75
  ) +
  scale_color_manual(
    values = c(
      "Tree-ring Density" = "#d95f02",
      "Radiocarbon Density" = "#1b9e77",
      "VEP Population" = "black"
    ),
    limits = c(
      "Tree-ring Density",
      "Radiocarbon Density",
      "VEP Population"
    )
  ) +
  xlab("Year AD") +
  scale_x_continuous(
    limits = c(200, 1800),
    breaks = seq(200, 1800, 200),
    expand = expansion(0, 0)
  ) +
  theme_minimal(18) +
  theme(
    legend.title = element_blank(),
    legend.justification = c(0, 1),
    legend.position = "bottom",
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(size = 10),
    axis.title.y = element_blank(),
    # axis.text.y = element_blank(),
    # panel.grid.major.y = element_blank(),
    # panel.grid.minor.y = element_blank(),
    plot.margin = unit(c(0.25, 0.5, 0.25, 0.25), "in")
  ) +
  facet_wrap(
    vars(`Study Area`),
    ncol = 1,
    scales = "free_y"
  )
```
